---
title: "J-COMP"
author: "MUSKAAN YADAV 20BCE1622"
date: "2022-11-05"
output: html_document
---

```{r}
library(rvest)
library(dplyr)
library(kableExtra) 
library(tidyverse)
library(rsoi)
library(plotly)
```

#Scraping from wikipedia
```{r}
url  <- "https://en.wikipedia.org/wiki/Accumulated_cyclone_energy"
# Read HTML page into wbpg object
wbpg <- read_html(url)
# Read table information from wbpg object into tbl object
tbl  <- html_nodes(wbpg, "table")
# Read required table and save to ACE_Yearly_DF data.frame
ACE_Yearly_DF <- html_table(tbl[grep("Accumulated Cyclone Energy - Atlantic\n",tbl,ignore.case = T)],fill = T)[[1]]
colnames(ACE_Yearly_DF)[1]<-"Year"
# Add Minor Tropical Storms column by calculating it from Tropical Storms, Hurricanes & Major Hurricanes
ACE_Yearly_DF$MTS <- ACE_Yearly_DF$TS-ACE_Yearly_DF$HU--ACE_Yearly_DF$MH
# Storing dataframe to csv 
write.csv(ACE_Yearly_DF,'ACE_Yearly_DF.csv')
```


```{r}
ACE_Yearly_DF <- read.csv('ACE_Yearly_DF.csv')
ACE_Yearly_DF

```
```{r}
# From 2000 to 2022 
ACE_Yearly_DF %>% slice(150:172) %>%
  kbl(caption = "ACE DataFrame (Yearly)") %>%
  kable_paper("hover", full_width = F, position = "center") %>% 
  row_spec(0,color="black",background = "yellow")
```

#Scrape monthly ACE data from the NOAA website
```{r}
wbpg2   <- readLines("https://www.psl.noaa.gov/gcos_wgsp/Timeseries/Hurricane/hurr.atl.ace.data")
# Eliminate first and last five lines
df.temp <- wbpg2[-c(1,171:178)]
# 
df.temp <- gsub(pattern = ' |  |   ',replacement = 'and',x = df.temp)
df.temp <- gsub(pattern = 'andand',replacement = 'and',x = df.temp)
df.temp <- str_replace(string = df.temp,pattern = 'and',replacement = '')
df.temp <- as.data.frame(df.temp)
clnames <- c("Year","Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
ACE_Monthly_DF <- df.temp %>% 
  separate(df.temp, clnames, "and") %>% 
  gather(key = "Month",value = "ACE",-Year) %>% 
  mutate(Year = as.numeric(Year), ACE = as.numeric(ACE)) %>% 
  arrange(Year) %>% 
  select(Year,Month,ACE)

write.csv(ACE_Monthly_DF,'ACE_Monthly_DF.csv')

```
#from 1910 to 1934 (ACE monthly data)
```{r}
ACE_Monthly_DF <- read.csv('ACE_Monthly_DF.csv')
ACE_Monthly_DF %>% slice(709:1000)%>%
  kbl(caption = "ACE DataFrame (Monthly)") %>%
  kable_paper("hover", full_width = F, position = "center") %>% 
  row_spec(0,color="black",background = "yellow")
```
#ONI data from RSOI package
```{r}
enso_monthly_DF <- download_enso()
enso_yearly_DF <- enso_monthly_DF %>% group_by(Year) %>% summarise(ONI=mean(ONI))
enso_yearly_DF <- enso_yearly_DF %>%  mutate(phase = case_when(ONI > 0.5 ~ "Warm Phase/El Nino", 
                                ONI < -0.5  ~ "Cool Phase/La Nina",
                                ONI < 0.5 & ONI >-0.5  ~ "Neutral Phase"))
write.csv(enso_monthly_DF,'enso_monthly_DF.csv')
write.csv(enso_yearly_DF,'enso_yearly_DF.csv')
enso_yearly_DF

```
#ONI monthly
```{r}
enso_monthly_DF=read.csv('enso_monthly_DF.csv')
enso_monthly_DF %>% slice(589:861)%>%
  kbl(caption = "ONI DataFrame (Monthly)") %>%
  kable_paper("hover", full_width = F, position = "center") %>% 
  row_spec(0,color="black",background = "yellow")


```
#ONI yearly
```{r}
enso_yearly_DF=read.csv("enso_yearly_DF.csv")
enso_yearly_DF %>% slice(50:72)%>%
  kbl(caption = "ONI DataFrame (yearly)") %>%
  kable_paper("hover", full_width = F, position = "center") %>% 
  row_spec(0,color="black",background = "yellow")



```


```{r}
# Merging ACE_Yearly_DF and enso_yearly_DF for ease. 
DF_Graph_Yearly <- merge(ACE_Yearly_DF,enso_yearly_DF,by="Year")

# Font Descriptions to use for Plot. 
t1 <- list(family = "Arial, sans-serif", size = 13, color = "black")
t2 <- list(family = "Arial, sans-serif", size = 12, color = "black")

#Creating a DF with Mean_ACE & Mean_Tropical Storms for use in plot 
hz_line_mean <- data.frame("Year"= DF_Graph_Yearly$Year,"Mean_TS"=rep(mean(ACE_Yearly_DF$TS),nrow(DF_Graph_Yearly)),"Mean_ACE"=rep(mean(ACE_Yearly_DF$ACE),nrow(DF_Graph_Yearly)))

fig.1 <- plot_ly(ACE_Yearly_DF)
# Adding Tropical Storms as bar chart
fig.1 <- fig.1 %>% add_trace(data=DF_Graph_Yearly,x=~Year,y=~MTS,type="bar",name="Minor Tropical Storms",marker = list(color = 'rgb(158,202,225)'))
# Adding Hurricanes as line chart
fig.1 <- fig.1 %>% add_trace(x=~Year, y=~HU, name = 'Hurricanes',type="scatter",mode="lines",line=list(width=4))
# Adding Major Hurricanes as line chart
fig.1 <- fig.1 %>% add_trace(x=~Year, y=~MH, name = 'Major Hurricanes',type="scatter",mode="lines",line=list(width=4))
#Adding Mean Tropical Storms Line to Plot
fig.1 <- fig.1 %>% add_trace(x=hz_line_mean$Year,y=hz_line_mean$Mean_TS,name="Mean Total Tropical Storms",            
                             type="scatter",mode="lines",line=list(width=3,dash='dash',color='blue'))
fig.1 <- fig.1 %>% layout(title="Tropical Storms - Atlantic Basin",
                          yaxis = list(title = 'Count of Tropical Storms & Hurricanes',titlefont=t1,tickfont=t2,showgrid=F,showline = T), 
                          annotations = list(text = "Increase in ACE since 2005", font=t1, x = 2010, y = 20,showarrow=FALSE ),
                          shapes = list(list(type = "rect", fillcolor = "grey", line = list(color = "grey"), opacity = 0.25,
                                              x0 = 2000, x1 = 2020, xref = "x",
                                              y0 = 0, y1 = 30, yref = "y")),
                          barmode = 'group',
                          legend = list(orientation = 'v',x=0.05,y=0.9),
                          xaxis=list(autotick=FALSE, dtick=5,title="Year",titlefont=t1,tickfont=t2,showgrid=F))
fig.1
```
#Visualizing ACE data over the years.
```{r}
fig.2 <- plot_ly(ACE_Yearly_DF)
# Adding ACE as bar chart
fig.2 <- fig.2 %>% add_trace(data=DF_Graph_Yearly,x=~Year,y=~ACE,type="bar",name="Minor Tropical Storms",marker = list(color = 'rgb(158,202,225)'))
#Adding Mean Tropical Storms Line to Plot
fig.2 <- fig.2 %>% add_trace(x=hz_line_mean$Year,y=hz_line_mean$Mean_ACE,name="Mean ACE",            
                             type="scatter",mode="lines",line=list(width=3,dash='dash',color='blue'))
fig.2 <- fig.2 %>% layout(title="Accumulated Cyclone Energy - Atlantic Basin",
                          yaxis = list(title = 'ACE (10^4 kn^2)',titlefont=t1,tickfont=t2,showgrid=F,showline = T), 
                          annotations = list(text = "Increase in ACE since 2005", font=t1, x = 2010, y = 200,showarrow=FALSE ),
                          shapes = list(list(type = "rect", fillcolor = "grey", line = list(color = "grey"), opacity = 0.25,
                                              x0 = 2000, x1 = 2020, xref = "x",
                                              y0 = 0, y1 = 250, yref = "y")),
                          barmode = 'group',
                          legend = list(orientation = 'v',x=0.05,y=0.9),
                          xaxis=list(autotick=FALSE, dtick=5,title="Year",titlefont=t1,tickfont=t2,showgrid=F))
fig.2
```
#ONI data over the years,relationship between El Nino & La Nina conditions.
```{r}
#Creating a DF with El Nino & La Nina thresholds
hz_line_ONI <- data.frame("Year"= DF_Graph_Yearly$Year,"LaNina"=rep(-0.5,nrow(DF_Graph_Yearly)),"ElNino"=rep(0.5,nrow(DF_Graph_Yearly)))

# Font Descriptions to use for Plot - Annotations
f1 <- list( family = "Arial, sans-serif", size = 13, color = "red" )
f2 <- list( family = "Arial, sans-serif", size = 13, color = "blue")


fig.3 <- plot_ly(DF_Graph_Yearly)
# Adding ONI as bar chart - red for positive & blue for negative
fig.3 <- fig.3 %>% add_trace(data=DF_Graph_Yearly,x=~Year,y=~ONI,type="bar",name="ONI",color = ~ONI < 0, colors = c("orangered", "royalblue"))
# Adding La Nina Threshold
fig.3 <- fig.3 %>% add_trace(x=hz_line_ONI$Year,y=hz_line_ONI$LaNina,name="La Nina Threshold",            
                             type="scatter",mode="lines",line=list(width=3,dash='dash',color='blue'))
# Adding El Nino Threshold
fig.3 <- fig.3 %>% add_trace(x=hz_line_ONI$Year,y=hz_line_ONI$ElNino,name="El Nino Threshold",            
                             type="scatter",mode="lines",line=list(width=3,dash='dash',color='red'))
fig.3 <- fig.3 %>% layout(title="Oceanic Niño Index",
                          yaxis = list(title = 'ONI (°C)',titlefont=t1,tickfont=t2,showgrid=F,showline = T),
                          annotations = list(
                            list(text = "El Nino Threshold", font=f1, x = 2009, y = 0.60,showarrow=FALSE ),
                            list(text = "La Nina Threshold", font=f2, x = 1963, y = -0.60,showarrow=FALSE )),
                          showlegend = FALSE,
                          xaxis=list(autotick=FALSE, dtick=5,title="Year",titlefont=t1,tickfont=t2,showgrid=F))
fig.3
```
#ACE V/s ONI ,taking  monthly data from 2000 to 2020 as a first step.
```{r}

# Merging ACE_Yearly_DF and enso_yearly_DF for ease.
DF_Graph_Monthly <- merge(ACE_Monthly_DF,enso_monthly_DF,by=c("Year","Month"))

#margin descriptions
m = list(l = 35,r = 40,b = 20,t = 50,pad = 1)

fig.4 <- plot_ly(DF_Graph_Monthly)
# Adding ACE data to Plot
fig.4 <- fig.4 %>% add_trace(data=DF_Graph_Monthly,x=~Date,y=~ACE,type="bar",name="ACE-Monthly")
#Adding ONI data to Plot
fig.4 <- fig.4 %>% add_trace(data=DF_Graph_Monthly,x=~Date,y=~ONI,type="bar",name="ONI-Monthly", yaxis='y2',marker=list(opacity=0.4))
fig.4 <- fig.4 %>%  layout(title="Accumulated Cyclone Energy - Atlantic Basin (2000-2020)",
                           yaxis = list(title = 'ACE (10^4 kn^2)', titlefont=t1,tickfont=t2,showgrid=F,showline = T,side='left',zeroline=F,range=c(-180,180)), 
                           yaxis2 = list(title = 'ONI °C', titlefont=t1,tickfont=t2,showgrid=F,showline = T,side='right',overlaying = "y",zeroline=F,range=c(-3,3)), 
                          legend = list(orientation = 'v',x=0.05,y=0.9),
                          xaxis=list(autotick=T,title="Year",titlefont=t1,tickfont=t2,showgrid=F,showline = T,zeroline=F,
                                     range = c("2005-01-01", "2020-01-01")),
                          margin = m)
fig.4
```
#determining relationship between ACE and ONI
```{r}

fig.6 <- plot_ly(DF_Graph_Yearly)
fig.6 <- fig.6 %>% add_trace(data=DF_Graph_Yearly,x=~ACE,y=~ONI,type="scatter",color=~phase)
fig.6 <- fig.6 %>% layout(title="Crossplot: ACE vs ONI",
                          yaxis = list(title = 'ONI (°C)',titlefont=t1,tickfont=t2,showgrid=F,showline = T,zeroline=F),
                          legend = list(orientation = 'v',x=0.7,y=0.9),
                          xaxis=list(autotick=F, dtick=20,title="ACE (10^4 kn^2)",titlefont=t1,tickfont=t2,showline = T,showgrid=F,
                                     type="linear",range=c(0,300)))
fig.6
```

#Conditional Probability for Monthly Data
#ACE > 10 = “Active”
#ACE < 4 = “Not Active”
#ACE > 4 & <10 = “Moderate”
```{r}

DF_Graph_Monthly <- DF_Graph_Monthly %>%  mutate(activity = case_when(ACE > 10 ~ "Active", 
                                ACE < 4  ~ "Not Active",
                                ACE > 4 & ACE <10 ~ "Moderate"))
cond_prob_monthly <- prop.table(table(DF_Graph_Monthly$activity, DF_Graph_Monthly$phase), margin = 1)
cond_prob_monthly %>% 
  kbl(caption = "Joint and Conditional Probability Tables - Monthly Data",digits=2) %>%
   kable_paper("hover", full_width = T, position = "center") %>% 
  row_spec(0,color="black",background = "yellow")
```
#Inference: El Niño conditions favors stronger hurricane activity in the central and eastern Pacific basins, and suppresses it in the Atlantic basin. Conversely, La Niña suppresses hurricane activity in the central and eastern Pacific basins, and enhances it in the Atlantic basin Let’s look at this by looking at the conditional probabilities.



#Conditional Probability for Annual data 

```{r}
DF_Graph_Yearly <- merge(ACE_Yearly_DF,enso_yearly_DF,by=c("Year"))
cond_prob_yearly <- prop.table(table(DF_Graph_Yearly$Classification, DF_Graph_Yearly$phase), margin = 1)

cond_prob_yearly %>% 
  kbl(caption = "Joint and Conditional Probability Tables - Yearly Data",digits=2) %>%
   kable_paper("hover", full_width = T, position = "center") %>% 
  row_spec(0,color="black",background = "yellow")
```
#Inference: It is hard to see a high hurricane activity during the La Nina season, we can  see a low occurance of cyclone activity during El Nino conditions.



#regression equation
```{r}
DF_Graph_Yearly$phase <- as.factor(DF_Graph_Yearly$phase) 
DF_Graph_Yearly$phase <- relevel(DF_Graph_Yearly$phase, ref = "Neutral Phase")
lm_yearly <- lm(DF_Graph_Yearly$ACE~DF_Graph_Yearly$ONI+DF_Graph_Yearly$phase,data=DF_Graph_Yearly)
summary(lm_yearly)

```

#plotting  the model on the crossplot to see the fit.

```{r}
prediction_df <- data.frame(DF_Graph_Yearly$ONI,DF_Graph_Yearly$phase)

prediction_answer <- predict(lm_yearly,prediction_df)

fig.7 <- plot_ly(DF_Graph_Yearly)
fig.7 <- fig.7 %>% add_trace(data=DF_Graph_Yearly,x=~ACE,y=~ONI,type="scatter",color=~phase)
fig.7 <- fig.7 %>% add_trace(data=DF_Graph_Yearly,x=~prediction_answer,y=~ONI,type="scatter",mode="line",color=~phase)
fig.7 <- fig.7 %>% layout(title="Crossplot: ACE vs ONI",
                          yaxis = list(title = 'ONI (°C)',titlefont=t1,tickfont=t2,showgrid=F,showline = T,zeroline=F),
                          legend = list(orientation = 'v',x=0.7,y=0.9),
                          xaxis=list(autotick=F, dtick=20,title="ACE (10^4 kn^2)",titlefont=t1,tickfont=t2,showline = T,showgrid=F,
                                     type="linear",range=c(0,300)))
fig.7
```
#The fit does not look very good.



# cross-plot of residual vs. ONI to check for any heteroskedasticity(variance of error term or residual term varies widely)
```{r}
fig.8 <- plot_ly(DF_Graph_Yearly)
fig.8 <- fig.8 %>% add_trace(data=DF_Graph_Yearly,x=~ONI,y=lm_yearly$residuals,type="scatter")
fig.8 <- fig.8 %>% layout(title="Crossplot of ONI vs Residuals",
                          yaxis = list(title = 'Residual',titlefont=t1,tickfont=t2,showgrid=F,showline = T,zeroline=F),
                          xaxis=list(autotick=T,title="ONI (°C)",titlefont=t1,tickfont=t2,showline = T,showgrid=F,zeroline=F,
                                     type="linear"))
fig.8
```

#Histogram of residuals or error 
```{r}
fig.9 <- plot_ly(DF_Graph_Yearly)
fig.9 <- fig.9 %>% add_trace(x=lm_yearly$residuals,type="histogram")
fig.9 <- fig.9 %>% layout(title="Histogram of Residuals",
                          yaxis = list(title = 'Frequency',titlefont=t1,tickfont=t2,showgrid=F,showline = T,zeroline=F),
                          xaxis=list(autotick=T,title="Residual Value",titlefont=t1,tickfont=t2,showline = T,showgrid=F,zeroline=F,
                                     type="linear"))

fig.9
```
```{r}
#correaltion yearly ACE V/S ONI
x=cor(ACE_Yearly_DF[1:100,3],enso_yearly_DF[1:100,2],  method = "pearson", use = "complete.obs")
x[1]

```

#The residuals look fairly normally distributed from the above plots.

#We can conclude that ONI is not a strong predictor of Hurricane Activity.We can also infer that there is no relationship between the ACE and ONI on the basis of conditional probability.ACE can be still be used as a parameter to predict Hurricane Activity as compared to ONI.
```{r}

```

